\color{navyblue} Difference-in-Differences Estimation with Spatial Spillovers

### 2021-04-28 latex/paper/Spillover.tex
4/28 Working Paper Draft

\addbibresource{references.bib}

@@

    \begin{abstract}
        Empirical work often uses treatment variables defined by geographic boundaries. When researchers ignore the common problem that the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I decompose this bias in two parts. First, the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment. Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias as well as semi-parametrically estimate the spillovers themselves. Lastly, I extend this estimation strategy in an event-study framework following \citet{Callaway_Sant’Anna_2020} to allow for staggered treatment adoption. Then, I turn to an empirical application revisiting an analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. I highlight the importance of considering general equilibrium spillovers in the analysis of place-based policies when estimating the direct effect on targeted areas.
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Last, in Section \ref{sec:event_study}, I extend estimation of the direct effect and spillover effects of treatment into the event study framework by extending the work of \citet{Callaway_Sant’Anna_2020}. This allows for a very common setting in which treatment turns on for different units in different periods. I first show how to adjust their methods to control for spatial spillovers in estimation of treatment effects. Then, I show how to estimate average spillover effects. My paper is the first paper to study estimation of treatment effects in an event-study framework in the presence of spillover effects.

@@

There is a large literature on estimation of treatment effects in the presence of spillovers using a `partial identification' framework where units are in distinct treatment clusters and outcomes depend on the treatment status within the observation's cluster only.\footnote{ \citet{Angelucci_DiMaro_2016} provides an overview of estimation of treatment effects in the presence of ``within-group'' spillovers. Examples in the literature include: \citet{Halloran_Struchiner_1995} consider community-vaccine rates in epidemology; \citet{Miguel_Kremer_2004} consider deworming programs in Kenyan schools; \citet{Sobel_2006} considers interference in the Moving to Opportunity Program; and \citet{Angrist_2014} studies the context of school peer effects.} Estimation compares untis in the partially treated clusters with control units in completely untreated clusters which do not receive spillover effects. This allows standard difference-in-differences estimation of both the direct effect (treatment effect on the treated) and spillover effects (treatment effect on the untreated in the treated clusters).

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} focuses on estimation of the direct and spillover effects of treatment. I evaluate currently used solutions in the literature and present a set of recommendations to more robust estimation strategies. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}. I apply these methods in evaluating spillover effects of U.S. Community Health Centers in Section \ref{sec:chc}.

@@

Following the canonical difference-in-differences framework, there is a time $t_0$ where treatment turns on and remains on afterwards.\footnote{This framework is extended to staggered treatment adoption in Section \ref{sec:event_study}.} Potential outcomes for unit $i$ at time $t$ are a function of own treatment-status $D_i$ and, departing from the standard potential outcomes framework, of a function of the entire vector of treatment assignments $h(\vec{D}, i)$ where $\vec{D} \in \{0,1 \}^n$ denotes the $n$-dimensional vector of all unit treatments. The potential outcomes are denoted $Y_{i,t}(D_i, h(\vec{D}, i)$). The function $h(\vec{D}, i)$ is referred to as an `exposure mapping' and is a non-negative scalar- or vector- valued function. The exposure mapping measures the intensity at which unit $i$ is affected by spatial spillovers. When unit $i$ is sufficiently `far' away, it has no exposure to spatial spillovers and $h(\vec{D}, i) = \vec{0}$. The exposure mapping formalizes the SUTVA violation by summarizing how outcomes are affected by other unit's treatment assignment. To help better understand the exposure mapping function, I give three examples that are commonly used in the literature.

\begin{example}
    First, $h(\vec{D}, i)$ can be an indicator variable that equals one only if there is a treated unit within $\bar{d}$ miles of unit $i$.\footnote{Similarly a dummy for counties that share contiguous borders is commonly used.} Let $d(i,j)$ be a distance measure which tells the distance unit $i$ is from unit $j$.  In this case \begin{equation}\label{eq:h_within}
        h(\vec{D}, i) = \max_{i \neq j} D_j * 1[ d(i,j) < \bar{d} ] 
    \end{equation}
    This exposure mapping is a realistic specification when it is assumed that spillover effects do not decay over distance until $\bar{d}$ and the intensity of spillovers does not depend on the number of neighboring units treated. For example, this exposure mapping likely applies in the context of new library creation \citep{Berkes_Nencka_2020}. In this case the distance $\bar{d}$ would be the maximum distance people would travel to a nearby library. The benefits of access to a neighboring town library does not depend on whether people can access 1 or more nearby libraries, so the binary exposure mapping is a good approximation to spillovers in this context.  
\end{example}

\begin{example}
    Second, $h(\vec{D}, i)$ can be a function that equals the number of units treated within distance $\bar{d}$, i.e. \begin{equation}\label{eq:h_within_additive}
        h(\vec{D}, i) = \sum_{j} D_j * 1[ d(i,j) < \bar{d} ].
    \end{equation}
    This exposure mapping is no longer binary, so the intensity of spillovers depend on the number of nearby units treated. In the context of large store openings, this exposure mapping captures the additive nature of agglomeration economies. As more nearby counties receive new stores, the agglomeration forces increase (e.g. \citet{Basker_2005}).
\end{example}

@@

@@

With the potential outcomes defined, I now formalize what is meant by `spatial spillovers'. I define `spillover onto control units' as: \[
    Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}).
\] 
The spillover measures the difference in non-treated potential outcomes between being exposed at intensity $h(\vec{D}, i)$ and not being exposed. Then, the average spillover effect onto control units averages over potential heterogeneity in the effect size of spillovers and over heterogeneity in exposure intensity $h(\vec{D}, i)$: \[
    \tau_{\text{spill,control}} \equiv \mathbb{E} \left[ Y_{it}(0, h(\vec{D}, i)) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 0 \right].
\]
To emphasize, the average spillover effect onto conrol units averages over each control unit's exposure mapping. For example, assume the potential outcomes is additively linear in $D_i$ and $h(\vec{D}, i)$ and that the coefficient $\beta_{\text{spill,control}}$ measures the effect of $h(\vec{D}, i)$ on outcome $Y$ among control units. Then an individual control unit's spillover effect is $\beta_{\text{spill,control}} \ h(\vec{D}, i)$. The average spillover effect onto control unit would therefore be $\tau_{\text{spill,control}} = \beta_{\text{spill,control}} * \mathbb{E} \left[ h(\vec{D}, i)\right]$, i.e. the average over all control units exposure mapping. Similarly, we define the average spillover effect onto also treated units as: \[ 
    \tau_{\text{spill,treated}} \equiv \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0) \ \vert \ D_i = 1 \right].
\]

It is important to clarify what I am assuming is the estimand of interest researchers would like to estimate when using difference-in-differences. I assume that what the `average treatment effect' is trying to measure in this context is what I will call the `direct effect of treatment': \[
    \tau_{\text{direct}} = \mathbb{E} \left[ Y_{it}(1, \vec{0}) - Y_{it}(0, \vec{0}) \ \vert \ D_i = 1 \right],
\] 
which measures the effect of being treated in the absence of exposure to spillovers.

My definition of the direct effect of treatment differs from \citet{Sävje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \mathbb{E} \left[ Y_{it}(1, h(\vec{D}, i)) - Y_{it}(0, h(\vec{D}, i)) \ \vert \ D_i = 1 \right],
\] 
where the expectation is over individuals and their exposures. The difference between their treatment effect and mine becomes more clear by adding and subtracting terms and rearranging:
\begin{align*}
    &Y_{it}(1, h(\vec{D}, i)) - Y_{it}(0, h(\vec{D}, i)) \\
    &\quad = Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0) + Y_{it}(1, 0) - Y_{it}(0,0) + Y_{it}(0,0) - Y_{it}(0, h(\vec{D}, i)) \\
    &\quad = \underbrace{Y_{it}(1, 0) - Y_{it}(0,0)}_{\text{Direct Effect}} + \underbrace{Y_{it}(1, h(\vec{D}, i)) - Y_{it}(1, 0)}_{\text{Spillover on Treated}} - \underbrace{(Y_{it}(0, h(\vec{D}, i) - Y_{it}(0,0))}_{\text{Spillover on Control}} 
\end{align*}
The individual treatment effect \citet{Sävje_Aronow_Hudgens_2019} use is the direct effect of treatment plus the difference between spillover on treated units and spillover on control units.

I prefer my definition of the direct effect because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. For example, a county deciding whether to implement a policy will want to consider the treatment effect plus the difference in spillover effects at a particular level of exposure. The estimand proposed by \citet{Sävje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at specific exposure intensities. Estimates of the direct effect and spillover effect can be combined to estimate the ATE as defined by \citet{Sävje_Aronow_Hudgens_2019}.

@@

In this section, I identify the two sources of bias in difference-in-differences estimation of the direct effect of treatment. Researchers typically estimate the canonical two-way fixed effects model, 
\begin{equation}\label{eq:twfe}    
    y_{it} = \tau D_{it} + \mu_i + \mu_t + \epsilon_{it},
\end{equation}
where $D_{it} = D_i 1(t=1)$. The estimator $\hat{\tau}$ is a biased estimate for $\tau_{\text{direct}}$ in the presence of spillovers. To show this, I first present the equivalent to the parallel counterfactual trends assumption in the context of the new potential outcome framework.

\begin{assumption}[Parallel Counterfactual Trends]\label{assumption:parallel}
    \[ 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 1 \right] = 
        \mathbb{E}\left[ Y_{i1}(0, \vec{0}) - Y_{i0}(0, \vec{0}) \ \vert \ D_i = 0 \right]
    \]
\end{assumption}
This assumption states that in the absence of treatment and with zero exposure (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status. This generalizes to the classic parallel counterfactual trends when SUTVA is satisfied because then every unit has zero exposure.

@@

\begin{proposition}[Decomposition of Difference-in-Differences Estimate]\label{thm:bias}\ \\    
    If Assumption \ref{assumption:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \underbrace{\mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 1 \right] - \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 0 \right]}_{\text{Difference-in-Differences}} \\ 
        &= 
        \mathbb{E} \left[ Y_{i1}(1, \vec{0}) - Y_{i1}(0, \vec{0}) \mid D_i = 1 \right] + \mathbb{E} \left[ Y_{i1}(1, h(\vec{D}, i)) - Y_{i1}(1, \vec{0}) \mid D_i = 1 \right] \\
        &\quad - \mathbb{E} \left[ Y_{i1}(0, h(\vec{D}, i)) - Y_{i1}(0, \vec{0}) \mid D_i = 0 \right] \\
        &= \tau_{\text{direct}} + \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}
    \end{align*}
\end{proposition}

The intuition behind the biases are as follows. First, the change in outcomes among treated units combines the direct effect and the spillover from nearby treated units. Therefore the first difference adds the average spillover effect onto the treated units, $\tau_{\text{spill,treated}}$. Second, the change in outcomes among control units combines the parallel counterfactual trend with the average spillover effect onto control units. Since $\hat{\tau}$ is found by subtracting this second difference, we subtract the average spillover effect onto the control, $\tau_{\text{spill,control}}$.

@@

% ------------------------------------------------------------------------------
\subsection{Removing Bias}\label{sec:remove_bias}
% ------------------------------------------------------------------------------

Under mild conditions, it is possible to control directly in a regression for the two sources of spillover effects. The set of indicators that contains all control/treated units respectively will estimate the average spillover effect on control/treated units and remove these sources of bias. The assumptions required to do this is that the researcher needs to be able to identify the maximum distance for which units are affected by spillovers and that there be unexposed units remaining.

@@

    (ii) There are treated and control units such that $\min_{j: \ D_j = 1} d(i,j) > \bar{d}$.

@@

Part (i) of Assumption (\ref{assumption:local}) requires that spillovers are `local' in that units are no longer exposed to spillovers after the distance $\bar{d}$. Part (ii) of Assumption (\ref{assumption:local}) requires that there exists units outside of this distance $\bar{d}$ so that the counterfactual trend can be estimated. Let $S_{i}= 1(\min_{j: \ D_j = 1} d(i,j) > \bar{d})$ be an indicator for being within $\bar{d}$ miles of the closest treated unit after treatment.\footnote{In contexts with different measures of `distance', note that this result only requires an indicator that captures all affected units. For example, thinking about spillovers between industries, this requires the researcher to correctly identify which industries are affected by spillovers and which are not.}

@@

\begin{equation}\label{eq:twfe_with_ind}    
    y_{it} = \tau D_{it} + \tau_{\text{spill,control}} (1-D_{it}) S_{i} + \tau_{\text{spill,treated}} D_{it} S_{i} +  \mu_i + \mu_t + \epsilon_{it}.
\end{equation}

\begin{proposition}[Unbiased Estimate for Direct Effect]\label{thm:remove_bias}
    If Assumption \ref{assumption:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe_with_ind}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 1, S_{i} = 0 \right] - \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 0, S_{i} = 0 \right] \\ 
        &= \tau_{\text{direct}} 
    \end{align*}
\end{proposition}

The intuition behind this is that the indicator interacted with treatment status serve as unbiased estimates of the \textit{average} spillover effect onto treated and control units and therefore remove these terms from the direct effect estimate, $\hat{\tau}$. The proof uses standard Frisch-Waugh-Lovell style logic on a set of indicator variables and the fact that if $S_{it} = 0$ then $h(\vec{D}, i) = \vec{0}$.\footnote{The derivation follows very closely the work of \citet{Clarke_2019}.}

This proposition shows that a researcher does not need to know how the spillovers occur over space in order to have an unbiased estimate for treatment effects. An indicator for being `near' treatment ineracted with treatment status is enough to have an unbiased estimate for the treatment effect. A natural question is why wouldn't researchers make the distance really large? The problem with this is that control units with $S_i = 0$ identify the counterfactual trend and treated units with $S_i = 0$ identify the direct effect. Therefore leaving few units with $S_i = 0$ will yield more variable estimates. On the other hand, having units that experience spillovers with Therefore there is a bias-variance trade-off that should be balanced by researchers.

In this section, I analyze methods that have been used by researchers for estimation of the direct and spillover effects. In doing so, I provide practical guidance for the estimation of treatment effects with spillovers. After evaluating currently used strategies, I propose a method that improves upon commonly used methods.

% ------------------------------------------------------------------------------
\subsection{Removing Bias by Dropping Control Units}
% ------------------------------------------------------------------------------

A common approach in the literature is to drop control units adjacent to treated units to avoid the problems of spillovers in identification. Proposition \ref{thm:bias} and Assumption \ref{assumption:local} would suggest that this can be effective in removing (at least part of) the average spillover effect on control units. However, since researchers do not know how far the spillovers extend there is difficulty in implementing this method effectively. On the one hand, removal of too few units will keep some control units that are experiencing spillover effects which would leave the estimated treatment effect biased. On the other hand, removing too many units will leave few observations and therefore increase the variance of the estimates. There is therefore a bias-variance trade off inherent to this method that can be readily improved by the method proposed in Section \ref{sec:parameterize}.

In addition to difficulties in implementation, the removal of problematic control observations, will not remove spillover effects onto treated units. Therefore if treated units are affected by the treatment status of nearby units, then the spillover effect on treated untis which would remain as a source of bias. Appendix \ref{sec:monte_carlo} contains a set of Monte Carlo simulations that highlight the problems more rigorously.

% ------------------------------------------------------------------------------
\subsection{Parameterization of Spillovers}
\label{sec:parameterize}
% ------------------------------------------------------------------------------

>>>>

%\addbibresource{references.bib}
\hypersetup{pdftitle={Difference-in-Differences Estimation with Spatial Spillovers}, pdfauthor={Kyle Butts}}

\usepackage{dcolumn}
\newcolumntype{d}[1]{D..{#1}}
\newcommand{\mcc}[1]{\multicolumn{1}{c@{}}{#1}}
\newcommand{\mcl}[1]{\multicolumn{1}{l@{}}{#1}}
\newcommand{\mcr}[1]{\multicolumn{1}{r@{}}{#1}}

@@

    \begin{abstract}
        Empirical work often uses treatment variables defined by geographic boundaries. When researchers ignore the common problem that the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I decompose this bias in two parts (1) the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment and (2) changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias as well as semi-parametrically estimate the spillover effects themselves. I extend this estimation strategy in an event-study framework following \citet{Callaway_SantAnna_2020} to allow for staggered treatment adoption. To highlight the importance of spillover effects, I revisit the analysis of three place-based policies and show the importance of directly controlling for general equilibrium spillovers caused by place-based policies.: the Tennessee Valley Authority \citep{Kline_Moretti_2014}, Federal Empowerment Zones \citep{Neumark_Young_2019}, and Community Health Centers \citep{Bailey_Goodman_Bacon_2015}. 
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Last, in Section \ref{sec:event_study}, I extend estimation of the direct effect and spillover effects of treatment into the event study framework by extending the work of \citet{Callaway_SantAnna_2020}. This allows for a very common setting in which treatment turns on for different units in different periods. I first show how to adjust their methods to control for spatial spillovers in estimation of treatment effects. Then, I show how to estimate average spillover effects. My paper is the first paper to study estimation of treatment effects in an event-study framework in the presence of spillover effects.

@@

There is a large literature on estimation of treatment effects in the presence of spillovers using a `partial identification' framework where units are in distinct treatment clusters and outcomes depend on the treatment status within the observation's cluster only.\footnote{ \citet{Angelucci_DiMaro_2016} provides an overview of estimation of treatment effects in the presence of ``within-group'' spillovers. Examples in the literature include: \citet{Halloran_Struchiner_1995} consider community-vaccine rates in epidemology; \citet{Miguel_Kremer_2004} consider deworming programs in Kenyan schools; \citet{Sobel_2006} considers interference in the Moving to Opportunity Program; and \citet{Angrist_2014} studies the context of school peer effects.} Estimation compares units in the partially treated clusters with control units in completely untreated clusters which do not receive spillover effects. This allows standard difference-in-differences estimation of both the direct effect (treatment effect on the treated) and spillover effects (treatment effect on the untreated in the treated clusters).

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} focuses on estimation of the direct and spillover effects of treatment. I evaluate currently used solutions in the literature and present a set of recommendations to more robust estimation strategies. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_SantAnna_2020}. I apply these methods in evaluating spillover effects of U.S. Community Health Centers in Section \ref{sec:chc}.

@@

This section will present a potential outcome framework that formalizes spatial spillovers. Following the canonical difference-in-differences framework, there are T periods and treatment turns on for all treated units at some period $t_0$ and remain on afterwards.\footnote{This framework is extended to staggered treatment adoption in Section \ref{sec:event_study}.} Potential outcomes, denoted $Y_{i,t}(D_i, h(\vec{D},i))$, for unit $i$ at time $t$ are a function of own treatment-status $D_i$ and, departing from the standard potential outcomes framework, of a function of the entire vector of treatment assignments $h(\vec{D}, i)$ where $\vec{D} \in \{0,1 \}^n$ denotes the $n$-dimensional vector of all unit treatments. The function $h(\vec{D}, i)$ is referred to as an `exposure mapping' and is a non-negative scalar- or vector- valued function. The exposure mapping measures the intensity at which unit $i$ is affected by spatial spillovers. When unit $i$ is sufficiently `far' away, it has no exposure to spatial spillovers and $h(\vec{D}, i) = \vec{0}$. The exposure mapping formalizes the SUTVA violation by summarizing how outcomes are affected by other unit's treatment assignment. To help better understand the exposure mapping function, I give three examples that are commonly used in the literature.

\begin{example}
    First, $h(\vec{D}, i)$ can be an indicator variable that equals one only if there is a treated unit within $\bar{d}$ miles of unit $i$.\footnote{Similarly a dummy for counties that share contiguous borders is commonly used.} Let $d(i,j)$ be a distance measure which tells the distance unit $i$ is from unit $j$.  In this case \begin{equation}\label{eq:h_within}
        h(\vec{D}, i) = \max_{i \neq j} D_j * \one[ d(i,j) < \bar{d} ] 
    \end{equation}
    This exposure mapping is a realistic specification when it is assumed that spillover effects do not decay over distance until $\bar{d}$ and the intensity of spillovers does not depend on the number of neighboring units treated. For example, this exposure mapping likely applies in the context of new library creation \citep{Berkes_Nencka_2020}. In this case the distance $\bar{d}$ would be the maximum distance people would travel to a nearby library. The benefits of access to a neighboring town library does not depend on whether people can access 1 or more nearby libraries, so the binary exposure mapping is a good approximation to spillovers in this context.  
\end{example}

\begin{example}
    Second, $h(\vec{D}, i)$ can be a function that equals the number of units treated within distance $\bar{d}$, i.e. \begin{equation}\label{eq:h_within_additive}
        h(\vec{D}, i) = \sum_{j} D_j * \one[ d(i,j) < \bar{d} ].
    \end{equation}
    This exposure mapping is no longer binary, so the intensity of spillovers is additive in the number of nearby units treated. In the context of large store openings, this exposure mapping captures the additive nature of agglomeration economies. As more nearby counties receive new stores, the force of agglomeration and therefore the spillover effect increases (e.g. \citet{Basker_2005}).
\end{example}

@@

% The results of this paper allows estimation and controlling for spillovers without needing to know the functional form of the exposure mapping. 
% TODO: Finish

@@

With the potential outcomes defined, I define `spillover effects' as: \[
    Y_{i,t}(D_i, h(\vec{D}, i)) - Y_{i,t}(D_i, \vec{0}).
\] 
The spillover effect measures the difference in potential outcomes between being exposed at intensity $h(\vec{D}, i)$ and not being exposed. This effect can differ for either treated or control units as the magnitude or even the nature of spillovers might differ between treated and untreated units. Then, the average spillover effect onto control/treated units averages over potential heterogeneity in the effect size of spillovers and over heterogeneity in exposure intensity $h(\vec{D}, i)$: \[
    \tau_{\text{spill,control}} \equiv \expec{Y_{i,t}(D, h(\vec{D}, i)) - Y_{i,t}(D, \vec{0}) \ \vert \ D_i = D}.
\]
To emphasize, the average spillover effect onto conrol/treated units averages over each unit's exposure mapping which can be heterogeneous across units. For example, spillover effects can decay over distance as in Equation (\ref{eq:h_decay}) and the average spillover effect averages across different exposure intensities.

It is important to clarify what I am assuming is the estimand of interest researchers would like to estimate when using difference-in-differences because there are multiple potential estimands defined in the literature. I assume that what the `average treatment effect' is trying to measure in this context is what I will call the `direct effect of treatment': \[
    \tau_{\text{direct}} = \expec{Y_{i,t}(1, \vec{0}) - Y_{i,t}(0, \vec{0}) \ \vert \ D_i = 1},
\] 
which measures the effect of being treated in the absence of exposure to spillovers.

My definition of the direct effect of treatment differs from \citet{Sävje_Aronow_Hudgens_2019} where they define the average treatment effect as \[ 
    \expec{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \ \vert \ D_i = 1},
\] 
where the expectation is over individuals and over different levels of exposures. The difference between their treatment effect and mine becomes more clear by adding and subtracting terms and rearranging:
\begin{align*}
    &Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0) + Y_{i,t}(1, 0) - Y_{i,t}(0,0) + Y_{i,t}(0,0) - Y_{i,t}(0, h(\vec{D}, i)) \\
    &\quad\quad = \underbrace{Y_{i,t}(1, 0) - Y_{i,t}(0,0)}_{\text{Direct Effect}} + \underbrace{Y_{i,t}(1, h(\vec{D}, i)) - Y_{i,t}(1, 0)}_{\text{Spillover on Treated}} - \underbrace{(Y_{i,t}(0, h(\vec{D}, i)) - Y_{i,t}(0,0))}_{\text{Spillover on Control}} 
\end{align*}
The individual treatment effect \citet{Sävje_Aronow_Hudgens_2019} use is the direct effect of treatment plus the difference between spillover on treated units and spillover on control units.

I prefer my definition of the direct effect because it allows for seperate identification of the direct effect of treatment and the spillover effects themselves. The alternative estimand averages over exposure intensities and therefore is difficult to be used by a policy maker in a particular location. That is, a county deciding whether to implement a policy will want to consider the treatment effect plus the difference in spillover effects at \textit{a particular} exposure intensity. The estimand proposed by \citet{Sävje_Aronow_Hudgens_2019} returns the average of spillovers across levels of exposure and can not be used to predict spillovers at specific exposure intensities while my preferred method of identifying the three effects seperately can. Estimates of the direct effect and spillover effect can be later combined to estimate the average treatment effect as defined by \citet{Sävje_Aronow_Hudgens_2019}.

@@

In this section, I identify the two sources of bias in difference-in-differences estimation of the direct effect of treatment. Researchers typically estimate the canonical two-way fixed effects model, 
\begin{equation}\label{eq:twfe}    
    y_{i,t} = \tau D_{i,t} + \mu_i + \mu_t + \epsilon_{i,t},
\end{equation}
where $D_{i,t} = D_i \one(t = 1)$. The estimator $\hat{\tau}$ is a biased estimate for $\tau_{\text{direct}}$ in the presence of spillovers. To show this, I first present the equivalent to the parallel counterfactual trends assumption in the context of the new potential outcome framework.

\begin{assumption}[Parallel Counterfactual Trends]\label{assumption:parallel}
    \[ 
        \expec{\bar{Y}_{i,Post}(0, \vec{0}) - \bar{Y}_{i,Pre}(0, \vec{0}) \ \vert \ D_i = 1 } = 
        \expec{\bar{Y}_{i,Post}(0, \vec{0}) - \bar{Y}_{i,Pre}(0, \vec{0}) \ \vert \ D_i = 0 },
    \]
    where $\bar{Y}_{i,W}$ is the average outcome in either the post- or pre-periods.
\end{assumption}
This assumption states that in the absence of treatment and with zero exposure (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status. This generalizes to the classic parallel counterfactual trends when SUTVA is satisfied because then every unit has zero exposure.

@@

\begin{proposition}[Decomposition of Difference-in-Differences Estimate]\label{thm:bias}\ \\    
    If Assumption \ref{assumption:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \underbrace{\expec{\bar{Y}_{i,Post} - \bar{Y}_{i,Pre} \ \vert \ D_i = 1 } - \expec{\bar{Y}_{i,Post} - \bar{Y}_{i,Pre} \ \vert \ D_i = 0 }}_{\text{Difference-in-Differences}} \\ 
        &= 
        \expec{\bar{Y}_{i,Post}(1, \vec{0}) - \bar{Y}_{i,Post}(0, \vec{0}) \mid D_i = 1} + \expec{\bar{Y}_{i,Post}(1, h(\vec{D}, i)) - \bar{Y}_{i,Post}(1, \vec{0}) \mid D_i = 1} \\
        &\quad\quad - \expec{\bar{Y}_{i,Post}(0, h(\vec{D}, i)) - \bar{Y}_{i,Post}(0, \vec{0}) \mid D_i = 0} \\
        &= \tau_{\text{direct}} + \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}
    \end{align*}
\end{proposition}

The intuition behind the biases are as follows. First, the change in outcomes among treated units combines the direct effect of treatment with the spillover effects caused by nearby treated units. Therefore the first difference adds the average spillover effect onto the treated units, $\tau_{\text{spill,treated}}$. Second, the change in outcomes among control units combines the parallel counterfactual trend with the average spillover effect onto control units. Since $\hat{\tau}$ is found by subtracting the change in outcomes among the control units, we subtract the average spillover effect onto the control, $\tau_{\text{spill,control}}$.

@@

% ------------------------------------------------------------------------------
\subsection{Unbiased Estimation of Direct Effect}\label{sec:remove_bias}
% ------------------------------------------------------------------------------

Under mild conditions, it is possible to control directly in a regression for the two sources of spillover effects and to have unbiased estimates of the direct effect of treatment. A pair of indicators that contains all control/treated units respectively will estimate the average spillover effect on control/treated units and remove these sources of bias. The assumptions required to do this is that the researcher needs to be able to identify the maximum distance for which units are affected by spillovers and that there be unexposed units remaining.

@@

    (ii) There are treated and control units, $i$, such that $\min_{j: \ D_j = 1} d(i,j) > \bar{d}$.

@@

Part (i) of Assumption (\ref{assumption:local}) requires that spillovers are `local' in that units are no longer exposed to spillovers after some maximum distance $\bar{d}$. Part (ii) of Assumption (\ref{assumption:local}) requires that there exists control units with no exposure in order to estimate the counterfactual trend and treated units with no exposure in order to estimate the direct effect. Let $S_{i}= \one(\min_{j: \ D_j = 1} d(i,j) \leq \bar{d})$ be an indicator for being within $\bar{d}$ miles of the closest treated unit after treatment.

@@

\begin{equation}\label{eq:twfe_with_ind}    
    y_{i,t} = \tau D_{i,t} + \tau_{\text{spill,control}} (1-D_{i,t}) S_{i} + \tau_{\text{spill,treated}} D_{i,t} S_{i} +  \mu_i + \mu_t + \epsilon_{i,t}.
\end{equation}

\begin{proposition}[Unbiased Estimate for Direct Effect]\label{thm:remove_bias}
    If Assumption \ref{assumption:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe_with_ind}) is
    \begin{align*}
        \expec{\hat{\tau}} &= \expec{\bar{Y}_{i,Post} - \bar{Y}_{i,Pre} \ \vert \ D_i = 1, S_{i} = 0 } - \expec{\bar{Y}_{i,Post} - \bar{Y}_{i,Pre} \ \vert \ D_i = 0, S_{i} = 0 } \\ 
        &= \tau_{\text{direct}} 
    \end{align*}

    The expectation of the two spillover effect estimates are 
    \begin{align*}
        \expec{\hat{\tau}_{\text{spill,conrol}}} = \expec{ \bar{Y}_{i,Post}(0,h(\vec{D}, i)) - \bar{Y}_{i,Post}(0, \vec{0}) \ \vert \ D_i = 0, S_i = 1}
    \end{align*}
    and 
    \begin{align*}
        \expec{\hat{\tau}_{\text{spill,treated}}} = \expec{ \bar{Y}_{i,Post}(1,h(\vec{D}, i)) - \bar{Y}_{i,Post}(1, \vec{0}) \ \vert \ D_i = 1, S_i = 1}
    \end{align*}
\end{proposition}

The intuition behind this is that the indicator interacted with treatment status serve as unbiased estimates of the \textit{average} spillover effect onto treated and control units and therefore remove these terms from the direct effect estimate, $\hat{\tau}$. The proof uses standard Frisch-Waugh-Lovell style logic on a set of indicator variables and the fact that if $S_{i} = 0$ then $h(\vec{D}, i) = \vec{0}$.\footnote{The derivation follows closely the work of \citet{Clarke_2019}.} This proposition shows that a researcher does not need to know how the spillovers occur over space in order to have an unbiased estimate for treatment effects. An indicator for being `near' treatment ineracted with treatment status is enough to have an unbiased estimate for the treatment effect.\footnote{In contexts with different measures of `distance', note that this result only requires an indicator that captures all affected units. For example, thinking about spillovers between industries, this requires the researcher to correctly identify which industries are affected by spillovers and which are not.}

A natural question is why wouldn't researchers make the distance really large in order to guarantee they have an unbiased estimate of $\tau_{\text{direct}}$? The problem with this is that control units with $S_i = 0$ identify the counterfactual trend and treated units with $S_i = 0$ identify the direct effect. Therefore leaving few units with $S_i = 0$ will yield more variable estimates. On the other hand, having units that experience spillovers not included in the $S_i$ indicator will leave some bias in the estimate. Although, since spillover effects typically will grow weaker over distance, the bias may not be particularly large if distant units are mistakenly treated as if they have zero exposure. Therefore there is a bias-variance trade-off in extending the extent of $S_i$ that should be balanced by researchers guided by their particular economic context.

A common approach in the literature is to drop control units adjacent to treated units to avoid the problems of spillovers in identification. Proposition \ref{thm:bias} and Assumption \ref{assumption:local} would suggest that this can be effective in removing (at least part of) the average spillover effect on control units. However, since researchers do not know how far the spillovers extend there is difficulty in implementing this method effectively. On the one hand, removal of too few units will keep some control units that are experiencing spillover effects which would leave the estimated treatment effect biased. On the other hand, removing too many units will leave few observations and therefore increase the variance of the estimates. There is therefore a bias-variance trade off inherent to this method that can be easily improved by creating an indicator for these units rather than dropping them from the sample. In addition to the unnecessary loss of precision induced by this method, the removal of problematic control observations, will not remove spillover effects onto treated units. Therefore if treated units are affected by the treatment status of nearby units, then the spillover effect on treated units which would remain as a source of bias.\footnote{Appendix \ref{sec:monte_carlo} contains a set of Monte Carlo simulations that highlight the problems more rigorously.}
